---
title: "Example: Testing Oracle compatible queries in the sql_test function"
author: "Jakob Skelmose, Rasmus Rask, Lars Nielsen, Charles Vesteghem, Jennifer Bartell, Martin Bøgsted"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Testing Oracle compatible queries in the sql_test function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(simulacrumR)
```

In  this example notebook, we want to demonstrate the cababilities of the sql_test function in the package, which is based on the sqldf library and which creates an temporal sqlite database in the R environment. 

For this demonstration, we use a guide for sql queries published by Health Data Insigt: "Simulacrum Guidance: Getting started with extracting data", which can be found on: https://simulacrum.healthdatainsight.org.uk/wp/wp-content/uploads/2018/11/SQL-Query-Guide.pdf. 


############################################################### 
############## Find the correct variable names 
############################################################### 

# Setup the database: 

``` {r}
dir <- "C:/Users/p90j/Desktop/Jakob/Data/Simulacrum/simulacrum_v2.1.0/Data/"
# Automated data loading 
data_frames_lists <- read_simulacrum(dir) 

SIM_AV_PATIENT <- data_frames_lists$sim_av_patient
SIM_AV_TUMOUR <- data_frames_lists$sim_av_tumour
SIM_AV_GENE <- data_frames_lists$sim_av_gene
SIM_RTDS_COMBINED <- data_frames_lists$sim_rtds_combined
SIM_RTDS_EPISODE <- data_frames_lists$sim_rtds_episode
SIM_RTDS_EXPOSURE <- data_frames_lists$sim_rtds_exposure
SIM_RTDS_PRESCRIPTION <- data_frames_lists$sim_rtds_prescription
SIM_SACT_CYCLE <- data_frames_lists$sim_sact_cycle
SIM_SACT_DRUG_DETAIL <- data_frames_lists$sim_sact_drug_detail
SIM_SACT_OUTCOME <- data_frames_lists$sim_sact_outcome
SIM_SACT_REGIMEN <- data_frames_lists$sim_sact_regimen
```


```{r}
query <- "SELECT *
FROM SIM_AV_PATIENT
INNER JOIN SIM_AV_TUMOUR ON SIM_AV_PATIENT.patientid = SIM_AV_TUMOUR.patientid;"


print(sql_test(query))
```


# 1. SELECT FROM 
```{r}
qeury <- "SELECT *
FROM SIM_AV_TUMOUR;"

print(sql_test(query))


query2 <- "SELECT diagnosisdatebest, site_icd10_o2
FROM SIM_AV_TUMOUR;"

print(sql_test(query2))
```
```{r}
summary(SIM_AV_TUMOUR)
```

# 2. WHERE
```{R}
query <- "SELECT *                          
FROM SIM_AV_TUMOUR
WHERE DIAGNOSISDATEBEST = 2014;
"

print(sql_test(query))


query2 <- "SELECT * 
FROM SIM_AV_TUMOUR
WHERE age>90;"

print(sql_test(query2))

query3 <- "SELECT * 
FROM SIM_AV_TUMOUR
WHERE age BETWEEN 18 AND 100;"

print(sql_test(query3))

query4 <- "SELECT * 
FROM SIM_AV_PATIENT
WHERE vitalstatusdate IS NULL;
"
print(sql_test(query4))

query5 <- "SELECT * 
FROM SIM_AV_TUMOUR
WHERE site_icd10_o2_3char IN ('C33','C34');"

print(sql_test(query5))

query6 <- "SELECT * 
FROM SIM_AV_TUMOUR
WHERE age = 34
AND gender = '2';" 

print(sql_test(query6))

query7 <- "SELECT COUNT (patientid) 
FROM SIM_AV_PATIENT
WHERE DEATHCAUSECODE_1A LIKE UPPER('%c50%');"

print(sql_test(query7))
```

# 3. Ordering 
``` {r}
query <- "SELECT * 
FROM SIM_AV_TUMOUR
ORDER BY diagnosisdatebest DESC;
"

print(sql_test(query))
```

# 4. Counting 
```{r}
query <- "SELECT COUNT(*)
FROM SIM_AV_TUMOUR;"

print(sql_test(query))

query2 <- "SELECT COUNT(*) 
FROM SIM_AV_PATIENT
WHERE vitalstatusdate IS NOT NULL;"

print(sql_test(query2))

query3 <- "SELECT COUNT (*) ,COUNT(tumourid) , COUNT(er_status) 
FROM SIM_AV_TUMOUR;"

print(sql_test(query3))
```

# 5. Counting in groups
```{r}
query <- "SELECT gender, COUNT(*) 
FROM SIM_AV_TUMOUR
GROUP BY GENDER
ORDER BY GENDER;"

print(sql_test(query))
```

# 6. Other aggregate functions (avg, min, max)
In query 2, 3 and 4 there is a example a sql function which convert interval or datetime into a datetime variable TO_CHAR 
``` {r}
query <- "
SELECT MAX(age), SUM(age), COUNT(age), AVG(age) 
FROM SIM_AV_TUMOUR;"

print(sql_test(query))

# query2 <- "SELECT DIAGNOSISDATEBEST, MAX(age), TO_CHAR (AVG(age),'9,999.99') 
# FROM SIM_AV_TUMOUR
# GROUP BY DIAGNOSISDATEBEST
# ORDER BY DIAGNOSISDATEBEST;"


#query3 <- "SELECT EXTRACT(YEAR FROM diagnosisdatebest) AS diagnosisyear, MAX(age), 
#TO_CHAR(AV(age),'9,999.99') 
#FROM SIM_AV_TUMOUR
#GROUP BY EXTRACT(YEAR FROM diagnosisdatebest)
#ORDER BY diagnosisyear;"


#query4 <- "SELECT diagnosisdatebest
#, TO_CHAR (diagnosisdatebest,'yyyy/mm/dd')
#, TO_CHAR (diagnosisdatebest,'dd/mm/yyyy')
#, TO_CHAR (diagnosisdatebest,'FMMonth DD, YYYY') 
#FROM SIM_AV_TUMOUR;"
```

# 7. Distinct 
In this markdown we get another example of the difference between sqlite and Oracle as we have a example where we want to extract Year 2014 from the diagnosisdatebest column. If we wanted to the same for sqlite, it should be written as in query instead of how it is written in query2 
``` {r}
query <- "
SELECT COUNT (DISTINCT patientid)
FROM SIM_AV_TUMOUR
WHERE site_icd10_o2_3char IN ('C50','C53','C54','C55','C56')
AND strftime('%Y', diagnosisdatebest) = '2014';"

print(sql_test(query))

query2 <- "SELECT COUNT (DISTINCT patientid), site_icd10_o2_3char
FROM SIM_AV_TUMOUR
WHERE site_icd10_o2_3char IN ('C50','C53','C54','C55','C56')
AND EXTRACT(YEAR FROM diagnosisdatebest)= 2014
GROUP BY site_icd10_o2_3char;"

print(sql_test(query2))
```

# 8. Case statements
Again does TO_CHAR and EXTRACT cause problems
``` {r}
query <- "SELECT age, 
CASE 
WHEN age <25 THEN 'Under 25' 
WHEN age <70 THEN '25 - 69'
ELSE '70+' 
END AS agegroup
FROM SIM_AV_TUMOUR;"

print(sql_test(query))

#query2 <- "SELECT patientid,
#CASE WHEN sex IN (2) THEN TO_CHAR(diagnosisdatebest, 'DD/MM/YYY')
#END AS diagdateiffemale
# FROM SIM_AV_TUMOUR
# WHERE site_icd10_o2_3char = 'C50' AND EXTRACT(YEAR FROM
#diagnosisdatebest)= 2014;"


#print(sql_test(query2))
```

# 9. Linking/joining tables 
``` {r}
query <- "SELECT *
FROM SIM_AV_PATIENT
LEFT OUTER JOIN SIM_AV_TUMOUR
ON SIM_AV_PATIENT.patientid = 
SIM_AV_TUMOUR.patientid;
"

print(sql_test(query))

query2 <- "SELECT *
FROM SIM_AV_PATIENT p
LEFT OUTER JOIN SIM_AV_TUMOUR t
ON p.patientid = t.patientid;"

print(sql_test(query2))


quey4 <- "SELECT tumourid, laterality
FROM SIM_AV_TUMOUR; "

print(sql_test(query4))
```

# 10. Subqueries 
These queries all contain either TO_CHAR or EXTRACT which means they cannot get executed in the sqlite 
```{r}
query <- "SELECT * 
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014;"

print(sql_test(query))


query2 <- "SELECT DISTINCT patientid
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014;"

print(sql_test(query2))

query3 <- "SELECT patientid, COUNT(*)
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014
GROUP BY patientid;"

print(sql_test(query3))

query4 <- "WITH breastpatients AS
(SELECT patientid, count(*) 
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014
GROUP BY patientid)
SELECT * FROM breastpatients;"

print(sql_test(query4))

query5 <- "WITH breastpatients AS
(SELECT patientid, count(*) 
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014
GROUP BY patientid)
SELECT COUNT(*)
FROM breastpatients;"

print(sql_test(query5))

query6 <- "WITH breastpatients AS
(SELECT patientid, count(*) as tumourcount
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014
GROUP BY patientid)SELECT tumourcount, count(*) 
FROM breastpatients
GROUP BY tumourcount
ORDER BY tumourcount asc;"


print(sql_test(query6))

query7 <- "SELECT tumourcount, count(*) 
FROM 
(SELECT patientid, count(*) as tumourcount
FROM sim_av_tumour
WHERE site_icd10_o2_3char = 'C50'
AND EXTRACT(YEAR FROM diagnosisdatebest) = 2014
GROUP BY patientid) breastpatients
GROUP BY tumourcount
ORDER BY tumourcount asc;"

print(sql_test(query6))
```
























